---
title: "How foreign aid is associated with economic growth?"
author: "Soutonnoma Sonia Nikiema"
date: "12/06/2021"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)  # ggplot(), %>%, mutate(), and friends
library(broom)  # Convert models to data frames
library(scales)  # Format numbers with functions like comma(), percent(), and dollar()
library(modelsummary)  # Create side-by-side regression tables
library(lubridate)
library(kableExtra)
library(MASS)
```



```{r include=FALSE}
list<-read.csv("data/list.csv", header= TRUE, check.names = FALSE)
rawODA<-read.csv("data/ODAP.csv", header= TRUE, check.names = FALSE)
rawGDP<-read.csv("data/GDP.csv", header= TRUE, check.names = FALSE)
rawdeflator<-read.csv("data/gdp_deflator.csv", header= TRUE, check.names = FALSE)
rawCapitalForm<-read.csv("data/capitalP.csv", header= TRUE, check.names = FALSE)
rawtrade<-read.csv("data/tradeP.csv", header= TRUE, check.names = FALSE)
rawunemployment<-read.csv("data/unemployment.csv", header= TRUE, check.names = FALSE)
rawpov<-read.csv("data2/pauvHeadcount.csv", header= TRUE, check.names = FALSE)

```






```{r include=FALSE}
rawODA <-rawODA%>%
    pivot_longer(!c('Country Name', 'Country Code'), names_to="years", values_to = "Aid", values_drop_na = TRUE)

rawGDP <-rawGDP%>%
    pivot_longer(!c('Country Name', 'Country Code'), names_to="years", values_to = "GDP", values_drop_na = TRUE)

rawdeflator <-rawdeflator%>%
    pivot_longer(!c('Country Name', 'Country Code'), names_to="years", values_to = "deflator", values_drop_na = TRUE)

rawtrade <-rawtrade%>%
    pivot_longer(!c('Country Name', 'Country Code'), names_to="years", values_to = "trade", values_drop_na = TRUE)

rawunemployment <-rawunemployment%>%
    pivot_longer(!c('Country Name', 'Country Code'), names_to="years", values_to = "U", values_drop_na = TRUE)

rawcapital <-rawCapitalForm%>%
    pivot_longer(!c('Country Name', 'Country Code'), names_to="years", values_to = "capital", values_drop_na = TRUE)

```






```{r include=FALSE}
SSAAid <- rawODA%>%
  rename(Country.Name='Country Name', Country.Code='Country Code')%>%
  filter(Country.Name %in% c("Angola", "Benin", "Botswana", "Burkina Faso", "Burundi", "Cape Verde", "Cameroon", "Central African Republic", "Chad", "Comoros", "Democratic Republic of the Congo", "Republic of the Congo", "Cote d'Ivoire", "Djibouti", "Equatorial Guinea", "Eritrea", "Ethiopia", "Gabon", "Gambia", "Ghana", "Guinea", "Guinea-Bissau", "Kenya", "Lesotho", "Liberia", "Madagascar", "Malawi", "Mali", "Mauritania", "Mauritius", "Mozambique", "Namibia", "Niger", "Nigeria", "Rwanda", "Sao Tome and Principe", "Senegal", "Seychelles", "Sierra Leone", "Somalia", "South Africa", "South Sudan", "Sudan", "Tanzania", "Togo", "Uganda", "Zambia", "Zimbabwe"))

SSAGDP <- rawGDP%>%
  rename(Country.Name='Country Name', Country.Code='Country Code')%>%
  filter(Country.Name %in% c("Angola", "Benin", "Botswana", "Burkina Faso", "Burundi", "Cape Verde", "Cameroon", "Central African Republic", "Chad", "Comoros", "Democratic Republic of the Congo", "Republic of the Congo", "Cote d'Ivoire", "Djibouti", "Equatorial Guinea", "Eritrea", "Ethiopia", "Gabon", "Gambia", "Ghana", "Guinea", "Guinea-Bissau", "Kenya", "Lesotho", "Liberia", "Madagascar", "Malawi", "Mali", "Mauritania", "Mauritius", "Mozambique", "Namibia", "Niger", "Nigeria", "Rwanda", "Sao Tome and Principe", "Senegal", "Seychelles", "Sierra Leone", "Somalia", "South Africa", "South Sudan", "Sudan", "Tanzania", "Togo", "Uganda", "Zambia", "Zimbabwe"))

SSAdeflator <- rawdeflator%>%
  rename(Country.Name='Country Name', Country.Code='Country Code')%>%
  filter(Country.Name %in% c("Angola", "Benin", "Botswana", "Burkina Faso", "Burundi", "Cape Verde", "Cameroon", "Central African Republic", "Chad", "Comoros", "Democratic Republic of the Congo", "Republic of the Congo", "Cote d'Ivoire", "Djibouti", "Equatorial Guinea", "Eritrea", "Ethiopia", "Gabon", "Gambia", "Ghana", "Guinea", "Guinea-Bissau", "Kenya", "Lesotho", "Liberia", "Madagascar", "Malawi", "Mali", "Mauritania", "Mauritius", "Mozambique", "Namibia", "Niger", "Nigeria", "Rwanda", "Sao Tome and Principe", "Senegal", "Seychelles", "Sierra Leone", "Somalia", "South Africa", "South Sudan", "Sudan", "Tanzania", "Togo", "Uganda", "Zambia", "Zimbabwe"))

SSAtrade <- rawtrade%>%
  rename(Country.Name='Country Name', Country.Code='Country Code')%>%
  filter(Country.Name %in% c("Angola", "Benin", "Botswana", "Burkina Faso", "Burundi", "Cape Verde", "Cameroon", "Central African Republic", "Chad", "Comoros", "Democratic Republic of the Congo", "Republic of the Congo", "Cote d'Ivoire", "Djibouti", "Equatorial Guinea", "Eritrea", "Ethiopia", "Gabon", "Gambia", "Ghana", "Guinea", "Guinea-Bissau", "Kenya", "Lesotho", "Liberia", "Madagascar", "Malawi", "Mali", "Mauritania", "Mauritius", "Mozambique", "Namibia", "Niger", "Nigeria", "Rwanda", "Sao Tome and Principe", "Senegal", "Seychelles", "Sierra Leone", "Somalia", "South Africa", "South Sudan", "Sudan", "Tanzania", "Togo", "Uganda", "Zambia", "Zimbabwe"))

SSAunemployment <- rawunemployment%>%
  rename(Country.Name='Country Name', Country.Code='Country Code')%>%
  filter(Country.Name %in% c("Angola", "Benin", "Botswana", "Burkina Faso", "Burundi", "Cape Verde", "Cameroon", "Central African Republic", "Chad", "Comoros", "Democratic Republic of the Congo", "Republic of the Congo", "Cote d'Ivoire", "Djibouti", "Equatorial Guinea", "Eritrea", "Ethiopia", "Gabon", "Gambia", "Ghana", "Guinea", "Guinea-Bissau", "Kenya", "Lesotho", "Liberia", "Madagascar", "Malawi", "Mali", "Mauritania", "Mauritius", "Mozambique", "Namibia", "Niger", "Nigeria", "Rwanda", "Sao Tome and Principe", "Senegal", "Seychelles", "Sierra Leone", "Somalia", "South Africa", "South Sudan", "Sudan", "Tanzania", "Togo", "Uganda", "Zambia", "Zimbabwe"))


SSAcapital <- rawcapital%>%
  rename(Country.Name='Country Name', Country.Code='Country Code')%>%
  filter(Country.Name %in% c("Angola", "Benin", "Botswana", "Burkina Faso", "Burundi", "Cape Verde", "Cameroon", "Central African Republic", "Chad", "Comoros", "Democratic Republic of the Congo", "Republic of the Congo", "Cote d'Ivoire", "Djibouti", "Equatorial Guinea", "Eritrea", "Ethiopia", "Gabon", "Gambia", "Ghana", "Guinea", "Guinea-Bissau", "Kenya", "Lesotho", "Liberia", "Madagascar", "Malawi", "Mali", "Mauritania", "Mauritius", "Mozambique", "Namibia", "Niger", "Nigeria", "Rwanda", "Sao Tome and Principe", "Senegal", "Seychelles", "Sierra Leone", "Somalia", "South Africa", "South Sudan", "Sudan", "Tanzania", "Togo", "Uganda", "Zambia", "Zimbabwe"))


```




```{r include=FALSE}
SSAData<-inner_join(SSAGDP, SSAdeflator)

SSAData<-inner_join(SSAData, SSAAid)

SSAData<-inner_join(SSAData, SSAtrade)

SSAData<-inner_join(SSAData, SSAunemployment)

SSAData<-inner_join(SSAData, SSAcapital)

```



## Landlocked countries
```{r echo=TRUE}
landlocked<-c("Botswana", "Burkina Faso", "Burundi", "Central African Republic", "Chad", "Ethiopia", "Lesotho", "Malawi", "Mali", "Niger", "Rwanda", "South Sudan", "Uganda", "Zambia", "Zimbabwe")

print(landlocked, quote = FALSE, max.levels = NULL, color=NA)
```

## Clean data

```{r}

SSAData<-SSAData%>%
  mutate(geography=ifelse(Country.Name %in%landlocked, "landloked", "coastal"),
         landlocked=ifelse(Country.Name %in%landlocked, "1", "0"),
         GDP=GDP/deflator,
         Aid.L=lag(Aid, n=3L),
         GDP.L=lag(GDP, n=1L),
        capital.L=lag(capital, n=2L),
         U.L=lag(U, n=1L),
       trade.L=lag(trade, n=1L))

SSAData$years <- as.Date(as.character(SSAData$years), format = "%Y")
SSAData$years <- year(SSAData$years)

SSAData$landlocked <- as.numeric(as.character(SSAData$landlocked))
  
write_csv(SSAData, "data/SSAData.csv")
```

## coastal countries

```{r echo=TRUE}
coastacountries1<-SSAData%>%
    filter(geography=="coastal")%>%
  dplyr::select(Country.Name)

unique_coastal<-unique(coastacountries1$Country.Name)
print(unique_coastal, quote = FALSE, max.levels = NULL)
```


```{r include=FALSE}

library(sf)
world_map <- read_sf("data2/ne_110m_admin_0_countries/ne_110m_admin_0_countries.shp")

Africa_list<-list$name

Africa1<-world_map%>%
  filter(NAME_LONG%in%Africa_list)

Africa<-world_map%>%
  filter(NAME_LONG==c(landlocked, unique_coastal))

Africa_land <- world_map %>% 
  filter(NAME_LONG%in%landlocked)

Africa_coast <- world_map %>% 
  filter(NAME_LONG%in%unique_coastal)

#world_sans_antarctica <- world_map %>% 
 # filter(ISO_A3 != "ATA")

#world<-world_sans_antarctica%>% 
#  filter(ISO_A3 != "USA")
```

## map
```{r echo=TRUE}
ggplot() +
  geom_sf(data = Africa1) +
  geom_sf(data = Africa_land, fill = "#EC8E55", color = "white") +
   geom_sf(data = Africa_coast, fill = "#192DA1", color = "white") +
    geom_sf_label(data = Africa_land, aes(label = NAME),
                nudge_y = 0.2)+
  theme_void()
```





## landlocked data summary

```{r echo=TRUE}

SumData_land<-SSAData%>%
  dplyr::select(GDP, Aid, capital, U, trade, geography)%>%
  filter(geography=="landloked")



Sum_land<-SumData_land%>%
  dplyr::select(GDP, Aid, capital, U, trade)%>%
 summarise_each(funs(obs =n(),
                     min = min, 
                      q25 = quantile(., 0.25), 
                      median = median,
                     mean=mean,
                      q75 = quantile(., 0.75), 
                      max = max, 
                      sd = sd))

Tidy_land <- Sum_land %>% 
  gather(stat, val) %>%
  separate(stat, into = c("var", "stat"), sep = "_") %>%
  spread(stat, val) %>%
  dplyr::select(var, obs, min, q25, median, mean, q75, max, sd) # reorder columns



fram_land<-as.data.frame(Tidy_land)

write.csv(fram_land, "data/fram_land.csv")
```

## coatal data summary

```{r echo=TRUE}
SumData_coast<-SSAData%>%
  dplyr::select(GDP, Aid, capital, U, trade, geography)%>%
  filter(geography=="coastal")



Sum_coast<-SumData_coast%>%
  dplyr::select(GDP, Aid, capital, U, trade)%>%
 summarise_each(funs(obs =n(),
                     min = min, 
                      q25 = quantile(., 0.25), 
                      median = median,
                     mean=mean,
                      q75 = quantile(., 0.75), 
                      max = max,
                      sd = sd))

Tidy_coast <- Sum_coast %>% 
  gather(stat, val) %>%
  separate(stat, into = c("var", "stat"), sep = "_") %>%
  spread(stat, val) %>%
  dplyr::select(var, obs, min, q25, median, mean, q75, max, sd) # reorder columns



fram_coast<-as.data.frame(Tidy_coast)

write.csv(fram_coast, "data/fram_coast.csv")

```





## Graph 1: Capital relative to GDP lagged 1 period for landlocked and coastal
```{r echo=TRUE}
SSAData_cap<-SSAData%>%
  dplyr::select(years, geography, capital.L)%>%
  group_by(years, geography)%>%
summarise(capital.L=mean(capital.L))

SSAData_cap%>%
  na.omit()%>%
  ggplot(aes(years, capital.L, colour=geography, group=geography))+
         geom_line(width=.5) +
  geom_point(size=3, shape=21, fill="white")+
  scale_color_manual(values = c("darkgreen", "red"))+
    labs(x = "Years", y = "% Capital")+
  theme_bw()
```

## Graph 2: Trade lagged 1 period relative to GDP for landlocked and coastal 

```{r echo=TRUE}
SSAData%>%
  na.omit()%>%
  dplyr::select(years, geography, trade.L)%>%
  group_by(years, geography)%>%
summarise(trade.L=mean(trade.L))%>%


  ggplot(aes(years, trade.L, colour=geography, group=geography))+
         geom_line(width=.5) +
  geom_point(size=3, shape=21, fill="white")+
  scale_color_manual(values = c("darkgreen", "red"))+
    labs(x = "Years", y = "% Trade")+
  theme_bw()
```

## Graph 3: Unemployment relative to the labor force lagged 1 period for landlocked and coastal 
```{r echo=TRUE}
SSAData%>%
  na.omit()%>%
  dplyr::select(years, geography, U.L)%>%
  group_by(years, geography)%>%
summarise(U.L=mean(U.L))%>%


ggplot(aes(years, U.L, colour=geography, group=geography))+
         geom_line(width=.5)+ 
  geom_point(size=3, shape=21, fill="white")+
  scale_color_manual(values = c("darkgreen", "red"))+
    labs(x = "Years", y = "% U")+
  theme_bw()
```




## Graph 4: Official Development Assistance(foreign aid) relative to GDP lagged 3 periods for landlocked and coastal 

```{r echo=TRUE}
SSAData%>%
  na.omit()%>%
  dplyr::select(years, geography, Aid.L)%>%
  group_by(years, geography)%>%
summarise(Aid.L=mean(Aid.L))%>%


  ggplot(aes(years, Aid.L, colour=geography, group=geography))+
         geom_line(width=.5) +
  geom_point(size=3, shape=21, fill="white")+
  scale_color_manual(values = c("darkgreen", "red"))+
    labs(x = "Years", y = "% Aid")+
  theme_bw()


```





## Graph 5: GDP between landlocked and coastal 

```{r echo=TRUE}
plot_data<-SSAData%>%
group_by(geography) %>%
  summarize(mean_GDP = mean(GDP),
            se_GDP = sd(GDP) / sqrt(n()),
            upper = mean_GDP + (1.96 * se_GDP),
            lower = mean_GDP + (-1.96 * se_GDP))



ggplot(plot_data, aes(x = geography, y = mean_GDP)) +
  geom_pointrange(aes(ymin = lower, ymax = upper, colour=geography),
                  color = "darkgreen", size = 1)+
  theme_bw()
```



## Model 1
 GDP = $\beta0$ + $\beta1$$Aid_i$ + $\beta2$$GDP_i(t-1)$ + $\beta3$$capital_i$ +$\beta4$$trade_i$ + $\beta5$$U_i$ + $\upsilon$
 

## Model 2

$GDP_it$ = $\beta_0$ + $\beta_1$$GDP_i(t-1)$ + $\beta_2$$Aid_i$ + $\beta_3$$capital_i$ +$\beta_4$$trade_i$ + $\beta_5$$U_i$ + $\delta_1$landlocked + $\delta_2$Aid.landlocked + $\upsilon$



## column 3 from result table 3
```{r echo=TRUE}
model_noLag <- lm(log(GDP) ~ log(GDP.L) + Aid  + capital + trade + U, data = SSAData)

summary(model_noLag)
```



```{r include=FALSE}

model_tot_No_Landlocked <- lm(log(GDP) ~ log(GDP.L) + Aid.L  + capital.L + trade.L + U.L, data = SSAData)

tidy(summary(model_tot_No_Landlocked, conf.int = TRUE))
```

```{r include=FALSE}
SSAData<-SSAData%>%
  mutate(Aid.Landlocked=landlocked*Aid)

model_total <- lm(log(GDP) ~ log(GDP.L) + Aid.L + capital.L + trade.L + U.L + landlocked + Aid.Landlocked, data = SSAData)

summary(model_total)
```

## column 4 from result table 3
```{r echo=TRUE}
SSAData<-SSAData%>%
  mutate(Aid.Landlocked=landlocked*Aid)

model_total_noLag <- lm(log(GDP) ~ log(GDP.L) + Aid + capital + trade + U+ landlocked + Aid.Landlocked, data = SSAData)

summary(model_total_noLag)
```



```{r include=FALSE}
dat_landlocked<-SSAData%>%
  filter(geography=="landloked")

model_land <- lm(log(GDP) ~ log(GDP.L) + Aid.L + capital.L + trade.L + U.L, data = dat_landlocked)

tidy(model_land, conf.int = TRUE)
```

## column 1 from result table 3
```{r}
dat_landlocked<-SSAData%>%
  filter(geography=="landloked")
model_land_noLag <- lm(log(GDP) ~ log(GDP.L) + Aid + capital + trade + U, data = dat_landlocked)

summary(model_land_noLag)
```




```{r include=FALSE}
dat_coast<-SSAData%>%
  filter(geography=="coastal")

model_coast <- lm(log(GDP) ~ log(GDP.L) + Aid.L + capital.L + trade.L + U.L, data = dat_coast)


tidy(model_coast, conf.int = TRUE)
```

## column 2 from result table 3
```{r}
dat_coast<-SSAData%>%
  filter(geography=="coastal")

model_coast_noLag <- lm(log(GDP) ~ log(GDP.L) + Aid + capital + trade + U, data = dat_coast)

summary(model_coast_noLag)
```




```{r include=FALSE}
#modelsummary(list("landlocked" = model_land, "coastal" = model_coast, "SSA"=model_tot_No_Landlocked, "SSA.Landlocked"=model_total), stars = TRUE)
```


## Result table 3
```{r}
modelsummary(list("landlocked" = model_land_noLag, "coastal" = model_coast_noLag, "SSA"=model_noLag, "SSA.Landlocked"=model_total_noLag), stars = TRUE)
```


















